Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-22 paper.tex
tweak S(C) Fig; finish its description

	Reduction \fbox{$\judgeSreduce{C}{s}{s}$}
	\[
	\begin{array}{rclr}
	\hiredruleS{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{ext(cast(T_1,l,T_2),\kappa)}
	}{}
	\redruleS{
		\sOOinspect{\eOOvar{x}}{\rho}{\kappa}
	}{	
		\sOOreturn{\rho(x)}{\kappa}
	}{}
	\redruleS{
		\sOOinspect{\eOOsole}{\rho}{\kappa}
	}{
		\sOOreturn{\hcvOOtt}{\kappa}
	}{}
	\hiredruleS{
		\sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\rho}{\kappa}
	}{
		\sOOreturn{(\hcvOOfun{id(T_1)}{\rho}{x}{e}{id(T_2)})}{\kappa}
	}{}
	\redruleS{
		\sOOinspect{(\eOOcons{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{\rho}{
			wrap(\kOOconsI{e_2}{\rho}{\kappa})
		}
	}{}
	\redruleS{
		\sOOinspect{(\eOOinl{e})}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{wrap(\kOOinl{\kappa})}
	}{}
	\redruleS{
		\sOOinspect{(\eOOinr{e})}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{wrap(\kOOinr{\kappa})}
	}{}
	\redruleS{
		\sOOinspect{(\eOOapp{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{\rho}{wrap(\kOOappI{e_2}{\rho}{\kappa})}}{}

@@

\funrule{
	\continue{v_1}{(\kOOconsI{e_2}{\rho}{\kappa})}}{
	\sOOinspect{e_2}{\rho}{wrap(\kOOconsII{v_1}{\kappa})}}{}

\hifunrule{
	\continue{v_2}{(\kOOconsII{v_1}{\kappa})}}{
	\sOOreturn{(\hcvOOcons{v_1}{id(T_1)}{v_2}{id(T_2)})}{\kappa}}{}

\hifunrule{
	\continue{v}{(\kOOinl{\kappa})}}{
	\sOOreturn{(\hcvOOinl{v}{id(T)})}{\kappa}}{}

\hifunrule{
	\continue{v}{(\kOOinr{\kappa})}}{
	\sOOreturn{(\hcvOOinr{v}{id(T)})}{\kappa}}{}

\funrule{
	\continue{v_1}{(\kOOappI{e_2}{\rho}{\kappa})}}{
	\sOOinspect{e_2}{\rho}{(\kOOappII{v_1}{\kappa})}}{}

%\funrule{
%	\continue{v_2}{(\kOOappII{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{\kappa})}}{
%applyCast(v_2,c_1) >>= \lambda v. \sOOinspect{e}{\rho[x:=v]}{ext(c,\kappa)}
%}{}

\hifunrule{
	\continue{v_2}{(\kOOappII{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{\kappa})}}{
\begin{cases}
	\sOOinspect{e}{\rho[x:=v']}{ext(c_2,\kappa)} \\
\;\;\;\;\;\sidecond{applyCast(v_2,c_1) = \rOOsucc{v'}}
\\
\sOOhalt{l} \\
\;\;\;\;\;\sidecond{applyCast(v_2,c_1) = \rOOfail{l}}
\end{cases}
}{}

@@

\funrule{
	\continue{v_2}{
		\kOOcaseII{v_1}{e_3}{\rho}{\kappa}
	}
}{
	\sOOinspect{e_3}{\rho}{
	\kOOcaseIII{v_1}{v_2}{\kappa}
	}
}{}

\hifunrule{
	\continue{v_3}{(\mathtt{case_3}\;(\hcvOOinl{v}{c})\;
		(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})
		\;\kappa)}
}{
	\sOOreturn{v}{wrap(\kOOappII{(\hcvOOfun{seq(c,c_1)}{\rho}{x}{e}{c_2})}{\kappa})}
}{}

\hifunrule{
	\continue{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}
	{(\mathtt{case_3}\;(\hcvOOinr{v}{c})\;v_2\;\kappa)}
}{
	\sOOreturn{v}{wrap(\kOOappII{(\hcvOOfun{seq(c,c_1)}{\rho}{x}{e}{c_2})}{\kappa})}
}{}

@@

\todoKC{describe transition rules}
We list a fraction of reduction rules due to space limitation.
\todo[inline]{Don't omit reduction rules, instead put some in the appendix. -JS
\\---\\They are all in the figure. I haven't updated this paragraph 
accordingly. -KC}

%When evaluating a cast expression, the machine moves the cast to the
%continuation and evaluates the inner expression.
%To apply a casted function, the machine first casts $v_1$, the
%operand, then applies the casted operand to $v_2$, the underlying
%function, and finally cast the result of the function application.
%%
%To take out the first (resp. second) part of a casted pair, the
%machine firstly take out the first (resp. second) part of $v$, the
%underlying pair, and cast the result.
%%
%%In case splitting, if the target value is a left injection, the
%%machine moves to a state that will apply the first continuation
%%function to the value inside the left injection. The case for right
%%injection is similar.
%%
%When the target value is a casted sum, the machine moves the cast from
%the target value to continuations functions.
%%
%To cast a value, the machine invokes $applyCast$ on the value.  If the
%cast succeeds, the machine returns the result to the next
%continuation.  If the cast fails, the machine halts with the blame
%label.

The transition relation $\judgeSreduce{C}{s}{s}$ uses functions provides by $C$ 
to represent cases and to apply casts.
%
When evaluating a cast expression, the machine extends the continuation with 
cast $cast(T_1,l,T_2)$. The function $ext(c,\kappa)$ composes $c$ with the cast 
at the top of $k$ by calling $seq$.
%
To construct a function, the machine fills the casts $id$.
%
To return a value to a continuation, the machine first applies the top cast to 
the value. If the application fails, the machine halts with the blame label 
from the failure. Otherwise, the machine handles the pre-continuation with 
$continue$.
%
When the machine constructs a pair, a left injection, or a right injection, it 
fills the casts with $id$ as well, just like how it did for functions. 
%
To apply a function, the machine applies the domain cast $c_1$ to the 
operand $v_2$, extends the continuation with the codomain cast $c_2$, and 
evaluate the function body.
%
To take out the first (resp. second) part of a pair ...
%
%%In case splitting, if the target value is a left injection, the
%%machine moves to a state that will apply the first continuation
%%function to the value inside the left injection. The case for right
%%injection is similar.
%%
%When the target value is a casted sum, the machine moves the cast from
%the target value to continuations functions.
To case spliting a variant, ...

%When values are constructed, their hypercoercion parts are filled with outputs 
%of $id$. For instance, when a function value is constructed, its first part 
%and last part are initialized to $id(T_1)$ and $id(T_2)$respectively.
%When evaluating a cast expression, the current continuation is extended with a 
%hypercoercion constructed by $cast$. $ext$ composes the new hypercoercion 
%with the hypercoercion on the top of the continuation.
%When evaluating a compound expression, the machine firstly construct the new 
%pre-continuation, then turn it to a continuation by adding an identity 
%hypercoercion at the top. For instance, when evaluating a \texttt{cons} 
%expression, the machine firstly construct $\hckOOconsI{e_2}{\rho}{\kappa}$, 
%the new pre-continuation, then call $cont$, which adds an identity cast to 
%form a continuation. 
%When a function call happens, the machine firstly cast the operand. If the 
%casting succeeds, the machine then evaluate the body in the extended 
%environment and the extended continuation. If the casting fails, the machine 
%then halts with the blame label.

>>>>

	Reduction \fbox{$\judgeSreduce{C}{s}{s}$}
	\[
	\begin{array}{rclr}
	\hiredruleS{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{ext(cast(T_1,l,T_2),\kappa)}
	}{}
	\redruleS{
		\sOOinspect{\eOOvar{x}}{\rho}{\kappa}
	}{	
		\sOOreturn{\rho(x)}{\kappa}
	}{}
	\redruleS{
		\sOOinspect{\eOOsole}{\rho}{\kappa}
	}{
		\sOOreturn{\hcvOOtt}{\kappa}
	}{}
	\hiredruleS{
		\sOOinspect{(\eOOlam{T_1}{T_2}{x}{e})}{\rho}{\kappa}
	}{
		\sOOreturn{(\hcvOOfun{id(T_1)}{\rho}{x}{e}{id(T_2)})}{\kappa}
	}{}
	\redruleS{
		\sOOinspect{(\eOOcons{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{\rho}{
			wrap(\kOOconsI{e_2}{\rho}{\kappa})
		}
	}{}
	\redruleS{
		\sOOinspect{(\eOOinl{e})}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{wrap(\kOOinl{\kappa})}
	}{}
	\redruleS{
		\sOOinspect{(\eOOinr{e})}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{wrap(\kOOinr{\kappa})}
	}{}
	\redruleS{
		\sOOinspect{(\eOOapp{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{\rho}{wrap(\kOOappI{e_2}{\rho}{\kappa})}}{}

@@

\funrule{
	\continue{v_1}{\kOOconsI{e_2}{\rho}{\kappa}}}{
	\sOOinspect{e_2}{\rho}{wrap(\kOOconsII{v_1}{\kappa})}}{}

\hifunrule{
	\continue{v_2}{\kOOconsII{v_1}{\kappa}}}{
	\sOOreturn{(\hcvOOcons{v_1}{id(T_1)}{v_2}{id(T_2)})}{\kappa}}{}

\hifunrule{
	\continue{v}{\kOOinl{\kappa}}}{
	\sOOreturn{(\hcvOOinl{v}{id(T)})}{\kappa}}{}

\hifunrule{
	\continue{v}{\kOOinr{\kappa}}}{
	\sOOreturn{(\hcvOOinr{v}{id(T)})}{\kappa}}{}

\funrule{
	\continue{v_1}{\kOOappI{e_2}{\rho}{\kappa}}}{
	\sOOinspect{e_2}{\rho}{
		wrap(\kOOappII{v_1}{\kappa})
	}}{}
\hifunrule{
	\continue{v_2}{(\kOOappII{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{\kappa})}}{
\begin{cases}
	\sOOinspect{e}{\rho[x:=v']}{ext(c_2,\kappa)} \\
\;\;\;\;\;\sidecond{applyCast(v_2,c_1) = \rOOsucc{v'}}
\\
\sOOhalt{l} \\
\;\;\;\;\;\sidecond{applyCast(v_2,c_1) = \rOOfail{l}}
\end{cases}
}{}

@@

\funrule{
	\continue{v_2}{
		\kOOcaseII{v_1}{e_3}{\rho}{\kappa}
	}
}{
	\sOOinspect{e_3}{\rho}{
	wrap(\kOOcaseIII{v_1}{v_2}{\kappa})
	}
}{}

\hifunrule{
	\continue{v_3}{
		\kOOcaseIII{(\hcvOOinl{v}{c})}{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{\kappa}
%		(\mathtt{case_3}\;()\;
%		()
%		\;)
	}
}{
	\sOOreturn{v}{wrap(\kOOappII{(\hcvOOfun{seq(c,c_1)}{\rho}{x}{e}{c_2})}{\kappa})}
}{}

\hifunrule{
	\continue{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}
	{
		\kOOcaseIII{(\hcvOOinr{v}{c})}{v_2}{\kappa}
	}
}{
	\sOOreturn{v}{wrap(\kOOappII{(\hcvOOfun{seq(c,c_1)}{\rho}{x}{e}{c_2})}{\kappa})}
}{}

@@

The transition relation $\judgeSreduce{C}{s}{s}$ uses functions provides by $C$ 
to represent cases and to apply casts.
%
When evaluating a cast expression, the machine extends the continuation with 
cast $cast(T_1,l,T_2)$. The function $ext(c,\kappa)$ composes $c$ with the cast 
at the top of $k$ by calling $seq$.
%
To construct a function, the machine fills the casts $id$.
%
To return a value to a continuation, the machine first applies the top cast to 
the value. If the application fails, the machine halts with the blame label 
from the failure. Otherwise, the machine handles the pre-continuation with 
$cont$.
%
When the machine constructs a pair, a left injection, or a right injection, it 
fills the casts with $id$ as well, just like how it did for functions. 
%
To apply a function, the machine applies the domain cast $c_1$ to the 
operand $v_2$, extends the continuation with the codomain cast $c_2$, and 
evaluate the function body.
%
To take out the first part of a pair, the machine returns the first value to a 
continuation extended with the first cast. Taking out the second part of a pair 
is similar.
%
%%In case splitting, if the target value is a left injection, the
%%machine moves to a state that will apply the first continuation
%%function to the value inside the left injection. The case for right
%%injection is similar.
%%
To case split a variant, the machine first look at the variant value. If the 
variant is a left injection, the machine composes the cast in the left 
injection with the domain casts of the first continuation function, then moves 
to a state that will apply the new function to the value in the left injection.
The case for right injections is similar.
